A spiral galaxy's mass distribution uncovered through lensing and dynamics

### 2015-11-13 docs/data_v3.01.tex
Included corrections in abstract and introduction.

\paragraph{Redshift.} [TO DO: WHO???] found from SDSS spectra that J1331 has two redshifts inside 1 arcsec: The smaller one, $z_d = 0.113$, is the redshift of J1331 itself, the larger one, $z_s = 0.254$, is the redshift of the lensed background source \citep{SWELLSIII}. According to its redshift and the WMAP5 [TO DO: Do I have to explain the Acronym???] cosmology \citep{WMAP5cosm}, J1331 has an angular diameter distance of 414 Mpc, which translates into a transverse scaling of 1 arcsec $\hat{=}$ 2.01 kpc.

\paragraph{HST imaging.} We use the follow-up HST imaging, with which \citet{SWELLSI} confirmed that J1331 is indeed a strong gravitational lens [TO DO: Was it actually SWELLS I that confirmed that???]. They performed high resolution imaging with the Hubble Space Telescope's (HST) Wide-Field Planetary Camera 2 (WFPC2) and its WF3 CCD chip.  The images we are using are a combination of four exposures with each an exposure of 400s and were drizzled to a pixel scale of 1 pixel = 0.05 arcsec. In particular we use the images in the filters F450W, to identify the positions of the bluish lensing images, and F814W to create a surface brightness MGE model of the reddish bulge (I-band), used in the JAM dynamical modelling.

\paragraph{Stellar kinematics.} For the dynamical modelling we also use the stellar kinematics for J1331 measured by \citet{SWELLSV}. They obtained long-slit spectra along J1331's major-axis with the Low Resolution Imaging Spectrograph (LRIS) on the Keck I 10m telescope. The width of the slot was 1 arcsec and the seeing Conditions had a FWHM of $\sim 1.1$ arcsec. Spectra for spatial bins of different widths along the major axis were extracted. Analogously to \citet{SWELLSII} they measured line-of-sight rotation velocities ($v_\text{rot}$) and stellar velocity dispersion ($\sigma$) by fitting Gaussian line profiles to emission lines in these spectra. Gas kinematics were extracted from fits to H$\alpha$ and NII lines, as tracers for inonized gas.
\\The stellar kinematics, $v_\text{rot}$, $\sigma$ and $v_\text{rms}^2=v_\text{rot}^2 + \sigma^2$ are shown in Figure \ref{fig:kinematics}. The rotation curve reveals a counter-rotating core within 2 arcsec $\simeq$ 4 kpc. Outside of $\sim 3.5$ arcsec there is a steep drop in the dispersion, which is exptected at the boundary between the pressure supported bulge and the rotationally supported disk, which appears around this radius in the F450W filter in Figure \ref{fig:F450W}. However, in the brighter F814W filter in Figure \ref{fig:F814W}  the large reddish bulge extends out to $\sim$5 arcsec. 
\\Inside of $\sim 4$ arcsec, the data appears to be symmetric, outside of this the assumption of axisymmetry seems not to be valid anymore, considering the data. We add -2.3 km/s to the $v_\text{rot}$ to ensure $v_\text{rot}(R'=0) \sim 0$ as a possible correction term for a systematic misjudgement of the systemic velocity. We also symmetrize the data within 4 arcsec and asign a minimum error of $\delta v_\text{rms} > 5$ km/s to the $v_\text{rms}$ data. In the JAM modelling, which is based on the assumption of axisymmetry, only kinematics within $\sim 2.5$ and 4 arcsec are used. Another reason to restrict to modelling on the bulge region is that our MGE in Table \ref{tab:MGEF814W} is only a good representation of J1331's F814W light distribution inside $\sim 5$ arcsec.\\

[TO DO: How to deal with the typo in J1331s name???? SWELLS call it J1331+3638, but its actual coordinates are 1331+3628.]

>>>>

\paragraph{Redshift.} \Wilma{[TO DO: WHO???]} found from SDSS spectra that J1331 has two redshifts inside $1''$: The smaller one, $z_d = 0.113$, is the redshift of J1331 itself, the larger one, $z_s = 0.254$, is the redshift of the lensed background source \citep{SWELLSIII}. According to its redshift and the WMAP5 \Wilma{[TO DO: Do I have to explain the Acronym???]} cosmology \citep{WMAP5cosm}, J1331 has an angular diameter distance of 414 Mpc, which translates into a transverse scaling of $1'' \hat{=} 2.01~\text{kpc}$.

\paragraph{HST imaging.} We use the follow-up HST imaging, with which \citet{SWELLSI} confirmed that J1331 is indeed a strong gravitational lens \Wilma{[TO DO: Was it actually SWELLS I that confirmed that???]}. They performed high resolution imaging with the Hubble Space Telescope's (HST) Wide-Field Planetary Camera 2 (WFPC2) and its WF3 CCD chip.  The images we are using are a combination of four exposures with each an exposure of 400s and were drizzled to a pixel scale of 1 pixel = $0.05''$. In particular we use the images in the filters F450W, to identify the positions of the bluish lensing images, and F814W to create a surface brightness MGE model of the reddish bulge (I-band), used in the JAM dynamical modelling.

\paragraph{Stellar kinematics.} For the dynamical modelling we also use the stellar kinematics for J1331 measured by \citet{SWELLSV}. They obtained long-slit spectra along J1331's major-axis with the Low Resolution Imaging Spectrograph (LRIS) on the Keck I 10m telescope. The width of the slot was $1''$ and the seeing conditions had a FWHM of $\sim 1.1''$. Spectra for spatial bins of different widths along the major axis were extracted. Analogously to \citet{SWELLSII} they measured line-of-sight rotation velocities ($v_\text{rot}$) and stellar velocity dispersion ($\sigma$) by fitting Gaussian line profiles to emission lines in these spectra. Gas kinematics were extracted from fits to H$\alpha$ and NII lines, as tracers for inonized gas.
\\The stellar kinematics, $v_\text{rot}$, $\sigma$ and $v_\text{rms}^2=v_\text{rot}^2 + \sigma^2$ are shown in Figure \ref{fig:kinematics}. The rotation curve reveals a counter-rotating core within $2''\simeq$ 4 kpc. Outside of $\sim 3.5''$ there is a steep drop in the dispersion, which is exptected at the boundary between the pressure supported bulge and the rotationally supported disk, which appears around this radius in the F450W filter in Figure \ref{fig:F450W}. However, in the brighter F814W filter in Figure \ref{fig:F814W}  the large reddish bulge extends out to $\sim5''$. 
\\Inside of $\sim 4''$, the data appears to be symmetric, outside of this the assumption of axisymmetry seems not to be valid anymore, considering the data. We add -2.3 km/s to the $v_\text{rot}$ to ensure $v_\text{rot}(R'=0) \sim 0$ as a possible correction term for a systematic misjudgement of the systemic velocity. We also symmetrize the data within $4''$ and asign a minimum error of $\delta v_\text{rms} > 5$ km/s to the $v_\text{rms}$ data. In the JAM modelling, which is based on the assumption of axisymmetry, only kinematics within $\sim 2.5$ and $4''$ are used. Another reason to restrict to modelling on the bulge region is that our MGE in Table \ref{tab:MGEF814W} is only a good representation of J1331's F814W light distribution inside $\sim 5''$.\\

\Wilma{[TO DO: How to deal with the typo in J1331s name???? SWELLS call it J1331+3638, but its actual coordinates are 1331+3628.]}

### 2015-11-13 docs/discussion_ML_v3.01.tex
Included corrections in abstract and introduction.

We compare our findings of the stellar mass-to-light ratios with the study by \cite{SWELLSV}: \cite{SWELLSV} found that the bulge of J1331 has an IMF more bottom-heavy than the Salpeter IMF from fitting a NFW halo to (1) the Einstein mass and (2) gas kinematics at larger radii $\gtrsim 8$ arcsec. In Section \ref{sec:results_JAM_NFW} we fitted the stellar kinematics within $\sim 3.5$ arcsec and the Einstein mass with a NFW halo and the best fit $\Upsilon_\text{I,*} = 4.2 \pm 0.2$ (see Table \ref{tab:modelB4_bestfit}) indicates a less bottom-heavy IMF than the Salpeter IMF. \cite{SWELLSV}  found systematically lower NFW halos ($v_\text{circ}(5\text{arcsec}) \sim 120 \text{km s}^{-1}$ according to their Figure 2) than we did ($v_\text{circ}(5\text{arcsec}) \sim 200 \text{km s}^{-1}$, see Figure \ref{fig:kinematics}). We note however, that our results would be much closer to those by \citet{SWELLSV}, if we only used the lensing results at small radii as they did: The total mass-to-light ratio of $\Upsilon_\text{I,tot}^\text{ein} = 5.6$ inside the Einstein radius (see Table \ref{tab:einsteinML}). In case of a lower mass dark matter halo, the largest contribution to this mass-to-light ratio would then be stellar mass. And the corresponding IMF would be more bottom-heavy than Salpeter,  like what \citet{SWELLSV} found. The circular velocity curve $v_\text{circ}(R) = \left. \sqrt{R \frac{\partial \Phi}{\partial R}} \right|_{z=0}$ \citet{SWELLSV} find for J1331 peaks at $v_\text{circ}(R\sim0.9 \text{arcsec}) \sim 350 \text{km s}^{-1}$ (from lensing constraints only) and at $v_\text{circ}(R\sim0.9 \text{arcsec}) \sim 370 \text{km s}^{-1}$ (when including a dark matter halo and gas kinematics in the fit). We get a very similar circular velocity curve (at least at small radii) for a mass-follows-light model with constant $\Upsilon_\text{I,tot}^\text{ein} = 5.6$, which peaks at $v_\text{circ}(R\sim0.9 \text{arcsec}) \sim 360 \text{km s}^{-1}$. From this comparison follows, that the model by \citet{SWELLSV} corresponds approximately to a mass-follows-light model in the inner regions of J1331. In Section \ref{sec:results_JAM_SB} however we showed, that this is \emph{not} a good model for J1331 and won't be able to reproduce the central stellar $v_\text{rms}$ dip. On the other hand, because our NFW JAM models fail to predict the stellar kinematics at large radii, the true halo parameters are more likely closer to the findings by \citet{SWELLSV}.\\

@@

Another feature in the stellar kinematics, that none of our JAM models was able to reproduce in the slightest and which we therefore excluded in the modelling, was the dip in $v_\text{rms}$ around $\sim 6$ arcsec, which occurs around the transition from bulge to disk (see Figures \ref{fig:F814W} and \ref{fig:kinematics}).\\

The above discussion motivates the following speculation: In the absence of a strong dark matter contribution in the center, the overall stellar-mass-to-light ratio within the Einstein radius indicates a bottom-heavy IMF close to the Salpeter IMF, consistent with estimates from the velocity dispersion. At the same time the central dip can then be only explained, if there was an increase in stellar mass-to-light ratio with radius \textit{within} the bulge. \emph{If} there was a central stellar population with an IMF close to the \citet{Chabrier2003} IMF, surrounded by a more bottom-heavy population, the central stellar kinematics would be well explained and be fully consistent with the lensing results. (According to Figure \ref{fig:lenscompareboth} the lensing result might not predict a mass to light ratio gradient inside $R_\text{ein}$, but then again the mass slope $alpha=1$ was only weakly constrained.) The disk of J1331 has a lower $\Upsilon_\text{I,*}$ than the bulge (see Table \ref{tab:previousresults} according to \citet{SWELLSI}). Such a drop in $\Upsilon_\text{I,*}$ at the transition region from bulge to disk around $\sim 5$ arcsec could lead to the observed drop in $v_\text{rms}$, while an increasing contribution of a lower mass dark matter halo at larger radii as found by \citet{SWELLSV}, would recover the kinematics in the out regions of J1331's disk.\\

[TO DO: Why are disk and bulge mass in SWELLS I different to the one in SWELLS IV??]

>>>>

We compare our findings of the stellar mass-to-light ratios with the study by \cite{SWELLSV}: \cite{SWELLSV} found that the bulge of J1331 has an IMF more bottom-heavy than the Salpeter IMF from fitting a NFW halo to (1) the Einstein mass and (2) gas kinematics at larger radii $\gtrsim 8''$. In Section \ref{sec:results_JAM_NFW} we fitted the stellar kinematics within $\sim 3.5''$ and the Einstein mass with a NFW halo and the best fit $\Upsilon_\text{I,*} = 4.2 \pm 0.2$ (see Table \ref{tab:modelB4_bestfit}) indicates a less bottom-heavy IMF than the Salpeter IMF. \cite{SWELLSV}  found systematically lower NFW halos ($v_\text{circ}(5'') \sim 120 \text{km s}^{-1}$ according to their Figure 2) than we did ($v_\text{circ}(5'') \sim 200 \text{km s}^{-1}$, see Figure \ref{fig:kinematics}). We note however, that our results would be much closer to those by \citet{SWELLSV}, if we only used the lensing results at small radii as they did: The total mass-to-light ratio of $\Upsilon_\text{I,tot}^\text{ein} = 5.6$ inside the Einstein radius (see Table \ref{tab:einsteinML}). In case of a lower mass dark matter halo, the largest contribution to this mass-to-light ratio would then be stellar mass. And the corresponding IMF would be more bottom-heavy than Salpeter,  like what \citet{SWELLSV} found. The circular velocity curve $v_\text{circ}(R) = \left. \sqrt{R \frac{\partial \Phi}{\partial R}} \right|_{z=0}$ \citet{SWELLSV} find for J1331 peaks at $v_\text{circ}(R\sim0.9'') \sim 350 \text{km s}^{-1}$ (from lensing constraints only) and at $v_\text{circ}(R\sim0.9'') \sim 370 \text{km s}^{-1}$ (when including a dark matter halo and gas kinematics in the fit). We get a very similar circular velocity curve (at least at small radii) for a mass-follows-light model with constant $\Upsilon_\text{I,tot}^\text{ein} = 5.6$, which peaks at $v_\text{circ}(R\sim0.9'') \sim 360 \text{km s}^{-1}$. From this comparison follows, that the model by \citet{SWELLSV} corresponds approximately to a mass-follows-light model in the inner regions of J1331. In Section \ref{sec:results_JAM_SB} however we showed, that this is \emph{not} a good model for J1331 and won't be able to reproduce the central stellar $v_\text{rms}$ dip. On the other hand, because our NFW JAM models fail to predict the stellar kinematics at large radii, the true halo parameters are more likely closer to the findings by \citet{SWELLSV}.\\

@@

Another feature in the stellar kinematics, that none of our JAM models was able to reproduce in the slightest and which we therefore excluded in the modelling, was the dip in $v_\text{rms}$ around $\sim 6''$, which occurs around the transition from bulge to disk (see Figures \ref{fig:F814W} and \ref{fig:kinematics}).\\

The above discussion motivates the following speculation: In the absence of a strong dark matter contribution in the center, the overall stellar-mass-to-light ratio within the Einstein radius indicates a bottom-heavy IMF close to the Salpeter IMF, consistent with estimates from the velocity dispersion. At the same time the central dip can then be only explained, if there was an increase in stellar mass-to-light ratio with radius \textit{within} the bulge. \emph{If} there was a central stellar population with an IMF close to the \citet{Chabrier2003} IMF, surrounded by a more bottom-heavy population, the central stellar kinematics would be well explained and be fully consistent with the lensing results. (According to Figure \ref{fig:lenscompareboth} the lensing result might not predict a mass to light ratio gradient inside $R_\text{ein}$, but then again the mass slope $alpha=1$ was only weakly constrained.) The disk of J1331 has a lower $\Upsilon_\text{I,*}$ than the bulge (see Table \ref{tab:previousresults} according to \citet{SWELLSI}). Such a drop in $\Upsilon_\text{I,*}$ at the transition region from bulge to disk around $\sim 5''$ could lead to the observed drop in $v_\text{rms}$, while an increasing contribution of a lower mass dark matter halo at larger radii as found by \citet{SWELLSV}, would recover the kinematics in the out regions of J1331's disk.\\

\Wilma{[TO DO: Why are disk and bulge mass in SWELLS I different to the one in SWELLS IV??]}

### 2015-11-13 docs/discussion_merger_v3.01.tex
Included corrections in abstract and introduction.

\paragraph{Counter-rotating core and minor merger.} J1331 has a large counter-rotating stellar core within $\sim 2$ arcsec. This suggests a process in J1331's past in which two components with angular momenta oriented in opposite directions were involved. Accretion of gas on retrograde orbits and subsequent star formation could lead to a younger and counter-rotating stellar population [TO DO: REF]. However, the to form enough stars such that the net rotation of the large and massive core is retrograde, a very large amount of gas would have had to be accreted by J1331- which is not very likely [TO DO: REF]. Another scenario are galaxy mergers. The products of major mergers of similarly massive disk galaxies are often quiescent, elliptical galaxies \citep{1972ApJ...178..623T} [TO DO: correct reference???]. Therefore a minor merger is more likely to have occured in J1331. The dense nucleus of a satellite galaxy on a retrograde orbit could survive the dissipationless accretion and could spiral to core due to tidal friction \citep{1984ApJ...287..577K}. Usually large bulges of spirals or ellipticals are reddest in the center and become bluer further out.  Because star formation in low-mass satellite galaxies occurred later than in massive galaxies [TO DO: REF], the merger remnant would now host a younger and bluer population with lower velocity dispersion within the older and redder bulge of the massive progenitor. Such a reverse color gradient could correspond to an increasing stellar mass-to-light ratio.
\\The minor merger theory could explain and support our findings from dynamical modelling and the previous sections, that J1331 might have an increasing stellar mass-to-light ratio within its core.

\paragraph{Possible modelling failures for merger remnants.} Another effect that minor mergers can have on galaxies are a misalignement (warp) of the photometric and kinematic major axis [TO DO: REF]. Would this be the case for J1331 the assumption of axisymmetry is not true anymore and our dynamical modelling would fail.
\\Mergers also could have changed the 3D shape of the dark matter halo and the NFW halo was therefore not a good model for J1331's dark matter halo. We also tried to model J1331 with a cored logarithmic halo. However, due to degeneracies in the modelling, we were not able to either constrain the profile for a cored logarithmic halo, or to rule it out.

>>>>

\paragraph{Counter-rotating core and minor merger.} J1331 has a large counter-rotating stellar core within $\sim 2''$. This suggests a process in J1331's past in which two components with angular momenta oriented in opposite directions were involved. Accretion of gas on retrograde orbits and subsequent star formation could lead to a younger and counter-rotating stellar population \Wilma{[TO DO: REF]}. However, the to form enough stars such that the net rotation of the large and massive core is retrograde, a very large amount of gas would have had to be accreted by J1331- which is not very likely [TO DO: REF]. Another scenario are galaxy mergers. The products of major mergers of similarly massive disk galaxies are often quiescent, elliptical galaxies \citep{1972ApJ...178..623T} [TO DO: correct reference???]. Therefore a minor merger is more likely to have occured in J1331. The dense nucleus of a satellite galaxy on a retrograde orbit could survive the dissipationless accretion and could spiral to core due to tidal friction \citep{1984ApJ...287..577K}. Usually large bulges of spirals or ellipticals are reddest in the center and become bluer further out.  Because star formation in low-mass satellite galaxies occurred later than in massive galaxies \Wilma{[TO DO: REF]}, the merger remnant would now host a younger and bluer population with lower velocity dispersion within the older and redder bulge of the massive progenitor. Such a reverse color gradient could correspond to an increasing stellar mass-to-light ratio.
\\The minor merger theory could explain and support our findings from dynamical modelling and the previous sections, that J1331 might have an increasing stellar mass-to-light ratio within its core.

\paragraph{Possible modelling failures for merger remnants.} Another effect that minor mergers can have on galaxies are a misalignement (warp) of the photometric and kinematic major axis \Wilma{[TO DO: REF]}. Would this be the case for J1331 the assumption of axisymmetry is not true anymore and our dynamical modelling would fail.
\\Mergers also could have changed the 3D shape of the dark matter halo and the NFW halo was therefore not a good model for J1331's dark matter halo. We also tried to model J1331 with a cored logarithmic halo. However, due to degeneracies in the modelling, we were not able to either constrain the profile for a cored logarithmic halo, or to rule it out.

### 2015-11-13 docs/introduction_v3.01.tex
Included corrections in abstract and introduction.

%Intro
Gravitational lensing and dynamical modelling provide independent constraints on the mass distribution of galaxies. Combining them allows for valuable cross-checking opportunities to resolve degeneracies inherent in both methods. It can also help to disentangle the stellar and dark matter content of galaxies, of which the latter is still shrouded by many open questions.\\

%Dark Matter
The flat rotation curves of galaxies (deduced from e.g. the stellar movements in the solar neighbourhood \citep{1932BAN.....6..249O} and $H\alpha$ rotation curves in external galaxies \citep{1978ApJ...225L.107R}) were the first indication that galaxies might reside in massive, almost spherical halos made of invisible ``dark'' matter. The standard model of cosmology, based on results by the Planck Mission \citep{WMAP5cosm}, predicts that $\sim 85\%$ of the universe's total matter content is in the form of non-baryonic cold dark matter. Simulations suggest that this kind of dark matter forms cuspy halos following a Navarro-Frenck-White (NFW) profile \citep{NFW96}. However, central dark matter density cusps are not observed in dark matter dominated dwarf galaxies [TO DO: REF]. If they exist in more massive galaxies depends strongly on stellar mass-to-light ratio [TO DO: REF]. Overall, observations suggest cored dark matter halos. This discrepancy, known as the core-cusp problem, might be due to a yet unknown interaction between dark matter and baryons (e.g. \citealt{2001ApJ...560..636E} [TO DO: more REF]).\\

Determining the overall mass distribution in massive galaxies and separating the dark from the stellar mass components is therefore a crucial step in better understanding the structure of galaxies and nature or dark matter.\\

%Lensing
Strong gravitational lensing is an independent method to tightly constrain the projected total mass of galaxies inside $\sim 1$ arcsec [TO DO: REF]. Massive galaxies can act as gravitational lenses and deflect the light of background sources, giving rise to multiple distorted images of the source [TO DO: REF]. By 2010 over 200 strong gravitational galaxy lenses had been discovered \citep{2010ARA&A..48...87T} and the number is still rising.\\

%Dynamical modelling
The mass profile at larger galactocentric radii can be probed by gas rotation curves that directly measure the galaxy's circlar velocity profile [TO DO: REF]. However, due to its dissipative nature, gas motions are very sensitive to disturbances by e.g. spiral arms and bars [TO DO: REF].\\
Very good and dissipationless dynamical tracers for the underlying gravitational potential are the stars. Their motions are complex, a bulk rotation around one principal axis combined with random motions in all coordinate directions, full dynamical modelling of rotation, dispersion and velocity anisotropies is needed to deduce the matter distribution. An example is axisymmetric Jeans modelling, which solves the Jeans equations for an assumed mass distribution and velocity anisotropy under the assumption of axisymmetry (e.g. \citealt{Cap08}). Degeneracies between stellar and dark matter and velocity anisotropy have to be resolved using additional assumptions.\\
Another dynamical modelling method often applied to external galaxies is e.g. Schwarzschild's orbital superposition approach \citep{schwarzschild}.\\
Because stars are present almost everywhere in the galaxy, stellar dynamical modelling can complement mass constraints from lensing at small and gas motions at large radii.\\

The Sloan WFC Edge-on Late-type Lens Survey (SWELLS, WFC = Wide field camera) \citep{SWELLSI,SWELLSII,SWELLSIII,SWELLSIV,SWELLSV,SWELLSVI} is dedicated to find and investigate spiral galaxies, seen almost edge-on so rotation curves can be easily measured, that are also strong gravitational lenses. By combining lensing and dynamical modelling degeneracies inherent in both methods can be broken. SWELLS picked galaxies in SDSS whose spectra had two different redshifts within 3 arcsec, which makes it likely to capture strong lenses with typical Einstein radii of 1 arcsec. Sufficiently inclined galaxies were picked by eye. Follow up high resolution imaging with the Hubble Space Telescope's (HST) Wide-Field Planetary Camera 2 (WFPC2) was performed to confirm their lensing status \citep{SWELLSI}.\\

%Characteristics of J1331
One of the SWELLS galaxies is SDSS J1331+3638, to which we refer as J1331 in the remainder of this work. It is a spiral galaxy of approximate Hubble type Sb at right ascension = 202.91800$^\circ$ and a declination = 36.46999$^\circ$ (epoch J2000). It has bluish spiral arms and a large reddish bulge (see Figures \ref{fig:F450W} and \ref{fig:F814W}), which is superimposed by quadruplet of extended bluish images approximately at a distance of 1 arcsec from the galaxy center (see Figure \ref{fig:lens_just_imgpos}). The lensed object might be a star-forming blob of a background galaxy. J1331 stands out of the SWELLS sample because of its large counter-rotating core (see Figure \ref{fig:kinematics}), which might be an indication that J1331 underwent a minor merger in its past.\\

@@

%Goal of this work
The goal of this work is to complement the previous work on J1331 by an in-depth analysis of the matter distribution in J1331's inner regions. We use stellar dynamical Jeans modelling in addition to lensing constraints, similar to a study of the Einstein Cross by \citet{GlennEC}. We attempt to disentangle the stellar and dark matter components and test if Jeans modelling works also in the presence of a counter-rotating core. Ideally this work on J1331 could also help understanding how minor mergers modify the mass distribution of a galaxy.\\

@@

\begin{figure*}
\centering
\begin{subfigure}{.5\textwidth}
  \centering
  \includegraphics[width=.9\linewidth]{fig/first_glimpse_450.ps}
  \caption{J1331 in F450W}
  \label{fig:F450W}
\end{subfigure}%
\begin{subfigure}{.5\textwidth}
  \centering
  \includegraphics[width=.9\linewidth]{fig/first_glimpse_814.ps}
  \caption{J1331 in F814W}
  \label{fig:F814W}
\end{subfigure}
\begin{subfigure}{.5\textwidth}
  \centering
  \includegraphics[width=.9\linewidth]{fig/lens_imgpos.ps}
  \caption{The lensing images}
  \label{fig:lens_just_imgpos}
\end{subfigure}%
\begin{subfigure}{.5\textwidth}
  \centering
  \includegraphics[width=.9\linewidth]{fig/stellar_kinematics_data.ps}
  \caption{Stellar Kinematics by \citet{SWELLSV}}
  \label{fig:kinematics}
\end{subfigure}
\caption{Hubble Space telescope (HST) images and stellar kinematics of the galaxy SDSS J1331+3638 (J1331), which has a large counter-rotating core and whose bulge acts as a strong lens for a bluish background source. \emph{Panel (a) and (b):} HST/WFPC2/WFC3 images of J1331 by \citet{SWELLSI} in two filters, F450W in panel (a) and F814W in panel (b). The galaxy's coordinates on the sky are right ascension $\alpha$ = 202.91800$^\circ$ and declination $\delta$ = 36.46999$^\circ$ (epoch J2000). Image orientation and scaling are indicated in panel (a); the scaling transformation from arcseconds to the physical size of the galaxy in kpc uses the galaxy's redshift $z_d = 0.113$ \citep{SWELLSIII} (i.e. assumes an angular diameter distance of 414 Mpc). The color scaling of these two images is the same. The black solid line in panel (b) shows the orientation of the major-axis. The line has a length of 10 arcsec and indicates the region within which we carry out the Jeans modelling. \emph{Panel (c):} The central region of J1331 in F450W, surface brightness subtracted. An IRAF ellipse [TO DO: How to write and reference????] fit to the F450W surface brightness in panel (a) was subtracted from the image. The (smoothed) residuals within the white square in panel (a) are shown in panel (c). Four bright blobs (A,B,C and D) become visible, which are arranged in a typical strong lensing configuration around the center of the galaxy (G). \emph{Panel (d):} Stellar Kinematics along the galaxy's major axis as measured by \citet{SWELLSV}, line-of-sight rotation velocity $v_\text{rot}$, line-of-sight velocity dispersion $\sigma$ and the rms-velocity $v_\text{rms} = \sqrt{v_\text{rot}^2 + \sigma^2}$. The dotted line in panel (b) indicates the galaxy's effective half-light radius (in the F814W filter), $R_\text{eff} = 2.6" = 5.2$ kpc. The $v_\text{rot}$ curve reveals that J1331 is counter-rotating within $R_\text{eff}$. [TO DO: Add (x,y) axis in figure b).???????]}
\label{fig:specialJ1331}
\end{figure*}

>>>>

%Intro
Gravitational lensing and dynamical modelling provide independent constraints on the mass distribution of galaxies. Combining them allows for valuable cross-checking opportunities to resolve degeneracies inherent in both methods. It can also help to disentangle the stellar and dark matter content of galaxies.

%Dark Matter
The flat rotation curves of galaxies \citep{1978ApJ...225L.107R} were the first indication that galaxies might reside in massive, almost spherical halos made of invisible ``dark'' matter. The standard model of cosmology, based on results by the Planck Mission \citep{WMAP5cosm} \Wilma{[TO DO: Aaron thinks Dunkley is a WMAP paper.... Check difference WMAP/Planck...]}, infers that $\sim 85\%$ of the universe's total matter content is in the form of non-baryonic cold dark matter. Simulations suggest that this kind of dark matter forms cuspy halos following a Navarro-Frenck-White (NFW) profile \citep{NFW96}. However, central dark matter density cusps are not observed in dark matter dominated dwarf galaxies \Wilma{[TO DO: REF: Aaron suggests Moore 1994, de Blok et al. 2001]}. Their existence in more massive galaxies depends strongly on stellar mass-to-light ratio \Wilma{[TO DO: REF: Aaron suggests Dutton et al. 2011 "dark halo response and the IMF"]}. Overall, observations suggest cored dark matter halos. This discrepancy, known as the core-cusp problem, might be resolved by galaxy formation processes such as mergers and outflows (e.g., \citealt{2001ApJ...560..636E} \Wilma{[TO DO: more REF: Aaron suggests Pontzen \& Governato 2012]}).

Determining the overall mass distribution in massive galaxies and separating the dark from the stellar mass components is therefore a crucial step in better understanding the structure of galaxies and nature or dark matter.

%Lensing
Strong gravitational lensing is an independent method to tightly constrain the projected total mass of galaxies inside $\sim 1''$ \Wilma{[TO DO: REF: Aaron suggests Treu 2010]}. Massive galaxies can act as gravitational lenses and deflect the light of background sources, giving rise to multiple distorted images of the source \Wilma{[TO DO: REF]}. By 2010 over 200 strong gravitational galaxy lenses had been discovered \citep{2010ARA&A..48...87T} and the number is still rising.

%Dynamical modelling
The mass profile at larger galactocentric radii can be probed by gas rotation curves that directly measure the galaxy's circlar velocity profile \Wilma{[TO DO: REF]}. However, due to its dissipative nature, gas motions are very sensitive to disturbances by e.g. spiral arms and bars \Wilma{[TO DO: REF]}.

Because stars are dissipationless dynamical tracers and present almost everywhere in the galaxy, stellar dynamical modelling can complement mass constraints from lensing at small and gas motions at large radii. But stellar motions are complex: a bulk rotation around one principal axis combined with random motions in all coordinate direction. Full dynamical modelling of rotation, dispersion and velocity anisotropies is needed to deduce the matter distribution.

The Sloan WFC Edge-on Late-type Lens Survey (SWELLS, WFC = Wide field camera) \citep{SWELLSI,SWELLSII,SWELLSIII,SWELLSIV,SWELLSV,SWELLSVI} is dedicated to find and investigate spiral galaxies, seen almost edge-on so rotation curves can be easily measured, that are also strong gravitational lenses. By combining lensing and dynamical modelling degeneracies inherent in both methods can be broken. SWELLS picked galaxies in SDSS whose spectra had two different redshifts within $3''$, which makes it likely to capture strong lenses with typical Einstein radii of $1''$. Sufficiently inclined galaxies were picked by eye. Follow up high resolution imaging with the Hubble Space Telescope's (HST) Wide-Field Planetary Camera 2 (WFPC2) was performed to confirm their lensing status \citep{SWELLSI}.

%Characteristics of J1331
One of the SWELLS galaxies is SDSS J1331+3638, to which we refer as J1331 in the remainder of this work. It is a spiral galaxy of approximate Hubble type Sb at right ascension = 202.91800$^\circ$ and a declination = 36.46999$^\circ$ (epoch J2000). It has bluish spiral arms and a large reddish bulge (see Figures \ref{fig:F450W} and \ref{fig:F814W}), which is superimposed by quadruplet of extended bluish images approximately at a distance of $1''$ from the galaxy center (see Figure \ref{fig:lens_just_imgpos}). The lensed object might be a star-forming blob of a background galaxy. J1331 stands out of the SWELLS sample because of its large counter-rotating core (see Figure \ref{fig:kinematics}), which might be an indication that J1331 underwent a minor merger in its recent past.

@@

%Goal of this work
The goal of this work is to complement the previous work on J1331 by an in-depth analysis of the matter distribution in J1331's inner regions. We use stellar dynamical modelling in addition to lensing constraints, similar to a study of the Einstein Cross by \citet{GlennEC}. We attempt to disentangle the stellar and dark matter components and test if employed axisymmetric Jeans models work also in the presence of a counter-rotating core. Ideally this work on J1331 could also help understanding how minor mergers modify the mass distribution of a galaxy.

@@

\begin{figure*}
\centering
\begin{subfigure}{.5\textwidth}
  \centering
  \includegraphics[width=.9\linewidth]{fig/first_glimpse_450.ps}
  \caption{J1331 in F450W}
  \label{fig:F450W}
\end{subfigure}%
\begin{subfigure}{.5\textwidth}
  \centering
  \includegraphics[width=.9\linewidth]{fig/first_glimpse_814.ps}
  \caption{J1331 in F814W}
  \label{fig:F814W}
\end{subfigure}
\begin{subfigure}{.5\textwidth}
  \centering
  \includegraphics[width=.9\linewidth]{fig/lens_imgpos.ps}
  \caption{The lensing images}
  \label{fig:lens_just_imgpos}
\end{subfigure}%
\begin{subfigure}{.5\textwidth}
  \centering
  \includegraphics[width=.9\linewidth]{fig/stellar_kinematics_data.ps}
  \caption{Stellar Kinematics by \citet{SWELLSV}}
  \label{fig:kinematics}
\end{subfigure}
\caption{Hubble Space telescope (HST) images and stellar kinematics of the galaxy SDSS J1331+3638 (J1331), which has a large counter-rotating core and whose bulge acts as a strong lens for a bluish background source. \emph{Panel (a) and (b):} HST/WFPC2/WFC3 images of J1331 by \citet{SWELLSI} in two filters, F450W in panel (a) and F814W in panel (b). The galaxy's coordinates on the sky are right ascension $\alpha$ = 202.91800$^\circ$ and declination $\delta$ = 36.46999$^\circ$ (epoch J2000). Image orientation and scaling are indicated in panel (a); the scaling transformation from arcseconds to the physical size of the galaxy in kpc uses the galaxy's redshift $z_d = 0.113$ \citep{SWELLSIII} (i.e. assumes an angular diameter distance of 414 Mpc). The color scaling of these two images is the same. The black solid line in panel (b) shows the orientation of the major-axis. The line has a length of $10''$ and indicates the region within which we carry out the Jeans modelling. \emph{Panel (c):} The central region of J1331 in F450W, surface brightness subtracted. An IRAF ellipse \Wilma{[TO DO: How to write and reference????]} fit to the F450W surface brightness in panel (a) was subtracted from the image. The (smoothed) residuals within the white square in panel (a) are shown in panel (c). Four bright blobs (A,B,C and D) become visible, which are arranged in a typical strong lensing configuration around the center of the galaxy (G). \emph{Panel (d):} Stellar Kinematics along the galaxy's major axis as measured by \citet{SWELLSV}, line-of-sight rotation velocity $v_\text{rot}$, line-of-sight velocity dispersion $\sigma$ and the rms-velocity $v_\text{rms} = \sqrt{v_\text{rot}^2 + \sigma^2}$. The dotted line in panel (b) indicates the galaxy's effective half-light radius (in the F814W filter), $R_\text{eff} = 2.6" = 5.2$ kpc. The $v_\text{rot}$ curve reveals that J1331 is counter-rotating within $R_\text{eff}$. \Wilma{[TO DO: Replace arcsec with '' in panel d)]} \Wilma{[TO DO: Mention, that the configuration of the additional two bright blobs do not suggest that they form a lensing doublet. They are probably background star formation regions as well. The blob above A could be stretched due to lensing and have a faint counter image between B and G. Maybe indicate in plot?]}}
\label{fig:specialJ1331}
\end{figure*}
